---
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Probability and Statistics

# Lab Assignment 1: Naive Bayes Classifier

## Work breakdown

-   Victor Muryn
-   Khrystyna Dmytriv
-   Veronika Tkachyshyn

## Introduction

During the first three weeks, you learned a couple of essential notions
and theorems, and one of the most important among them is the *Bayes
theorem*.

One of its applications is **Naive Bayes classifier**, which is a
probabilistic classifier whose aim is to determine which class some
observation probably belongs to by using the Bayes formula:
$$\mathsf{P}(\mathrm{class}\mid \mathrm{observation})=\frac{\mathsf{P}(\mathrm{observation}\mid\mathrm{class})\mathsf{P}(\mathrm{class})}{\mathsf{P}(\mathrm{observation})}$$

Under the strong independence assumption, one can calculate
$\mathsf{P}(\mathrm{observation} \mid \mathrm{class})$ as
$$\mathsf{P}(\mathrm{observation}) = \prod_{i=1}^{n} \mathsf{P}(\mathrm{feature}_i), \qquad \mathsf{P}(\mathrm{observation} \mid \mathrm{class}) = \prod_{i=1}^{n} \mathsf{P}(\mathrm{feature}_i \mid \mathrm{class}),$$
where $n$ is the total number of features describing a given
observation. Thus, $\mathsf{P}(\mathrm{class}|\mathrm{observation})$ now
can be calculated as

$$\mathsf{P}(\mathrm{class} \mid \mathrm{\mathrm{observation}}) = \mathsf{P}(\mathrm{class})\times \prod_{i=1}^{n}\frac{\mathsf{P}(\mathrm{feature}_i\mid \mathrm{class})}{\mathsf{P}(\mathrm{feature}_i)}\tag{1}$$

All the terms on the right-hand side can be estimated from the data as
respective relative frequencies;\
**\*see [this
site](https://monkeylearn.com/blog/practical-explanation-naive-bayes-classifier/)
for more detailed explanations\***

## Data description

There are 5 datasets uploaded on the CMS (data.zip)

To determine your variant, take your team number from the list of teams
on cms and take *mod 5* - this is the number of your data set.

-   **0 - authors** This data set consists of citations of three famous
    writers: Edgar Alan Poe, Mary Wollstonecraft Shelley and HP
    Lovecraft. The task with this data set is to classify a piece of
    text with the author who was more likely to write it.

Each data set consists of two files: *train.csv* and *test.csv*. The
first one you will need find the probabilities distributions for each of
the features, while the second one is needed for checking how well your
classifier works.

```{r}
# here goes a list of recommended libraries,
# though you may install other ones if they are needed
library(tidytext)
library(readr)
library(dplyr)
library(ggplot2)
library(comprehenr)

# for diagrams
library(wordcloud)
library(RColorBrewer)
library(wordcloud2)
library(tm)
library(stringr)
```

## Outline of the work

1.  **Data pre-processing** (includes removing punctuation marks and
    stop words, representing each message as a bag-of-words)
2.  **Data visualization** (it's time to plot your data!)
3.  **Classifier implementation** (using the training set, calculate all
    the conditional probabilities in formula (1) and then use those to
    predict classes for messages in the testing set)
4.  **Measurements of effectiveness of your classifier** (accuracy,
    precision and recall curves, F1 score metric etc)
5.  **Conclusions**

*!! do not forget to submit both the (compiled) Rmd source file and the
.html output !!*

## Data pre-processing

-   Read the *.csv* data files.
-   Сlear your data from punctuation or other unneeded symbols.
-   Clear you data from stop words. You don't want words as is, and, or
    etc. to affect your probabilities distributions, so it is a wise
    decision to get rid of them. Find list of stop words in the cms
    under the lab task.
-   Represent each test message as its bag-of-words. Here:
    <https://machinelearningmastery.com/gentle-introduction-bag-words-model/>
    you can find general introduction to the bag-of-words model and
    examples on to create it.
-   It is highly recommended to get familiar with R dataframes, it would
    make the work much easier to do.
-   Useful links:
    -   <https://steviep42.github.io/webscraping/book/bagofwords.html#tidytext> -
        example of using *tidytext* to count frequencies of the words.
    -   Basics of Text Mining in R:
        <http://rstudio-pubs-static.s3.amazonaws.com/256588_57b585da6c054349825cba46685d8464.html>
        . Note that it also includes an example on how to create a bag
        of words from your text document.

```{r}
test_path <- "data/0-authors/test.csv"
train_path <- "data/0-authors/train.csv"

stop_words <- read_file("stop_words.txt")
# https://stackoverflow.com/questions/27195912/why-does-strsplit-return-a-list
splitted_stop_words <- strsplit(stop_words, split = '\n')
splitted_stop_words <- splitted_stop_words[[1]]
```

```{r}
train <-  read.csv(file = train_path, stringsAsFactors = FALSE)
test <-  read.csv(file = test_path, stringsAsFactors = FALSE)
```

```{r}
# note the power functional features of R bring us! 
tidy_text <-
  unnest_tokens(train, 'splitted', 'text', token = "words") %>%
  filter(!splitted %in% splitted_stop_words)

unique_words = tidy_text %>% count(splitted)

```

## Data visualization

Each time you work with some data, you need to understand it before you
start processing it. R has very powerful tools to make nice plots and
visualization. Show what are the most common words for negative and
positive examples as a histogram, word cloud etc. Be creative!

```{r}
head(unique_words[order(unique_words$n, decreasing = TRUE), ], 10)
```

```{r}
set.seed(5433) # for reproducibility
dev.new(width = 1000,
        height = 1000,
        unit = "px")
wordcloud(
  words = unique_words$splitted,
  freq = unique_words$n,
  max.words = 1000,
  random.order = FALSE,
  rot.per = 0.5,
  scale = c(4, .5),
  min.freq = 165,
  colors = brewer.pal(8, "Dark2")
)
```

## Classifier implementation

```{r}
naiveBayes <- setRefClass(
  "naiveBayes",

  # attributes of classifier
  fields = list(
    count = "data.frame", # count of every word
    writers = "data.frame", # data about writers
    words = "numeric" # all number of words
  ),
  
  methods = list(
    # train our classifier
    fit = function(df) {
      .self$count <- df %>% count(author, splitted) # count number of words for each writer
      authors <- unique(count$author)
      sum_of_words <- sum(count$n)
      
      # dataset for writers
      writers_temp <- data.frame(
        author = character(0),
        probability = numeric(0),
        words = numeric(0)
      )
    
      # calculate data for each writer
      for (author in authors) {
        author_words <- sum(count$n[count$author == author])
        probability <- author_words / sum_of_words

        temp_df <- data.frame(
          author = author,
          probability = probability,
          words = author_words
        )

        writers_temp <- rbind(writers_temp, temp_df)
      }

      rownames(writers_temp) <- NULL
      .self$writers <- writers_temp
    
      # count words
      .self$words <- sum(.self$count$n)
    },


    # return prediction for a single message
    predict = function(message) {
      # clear income message
      message <- gsub('[\\(\\!\\)\\?\\,\\.]', '', message)
      message <- tolower(message)
      parts <- strsplit(message, split = " ")[[1]]
      parts <- parts[!parts %in% splitted_stop_words]

      # predicted result
      result_author <- ""
      max_probability <- 0

      for (i in 1:nrow(.self$writers)) {
        author <- .self$writers[i, ]
        
        # default probability
        # it's so huge, 'cause (0.5)^400 is very close to 0, that R rounds it to 0
        probability <- 1.797692e+308
        
        # for every word calculate probability of that word
        for (part in parts) {
          number <- .self$count$n[
            .self$count$author == author$author & .self$count$splitted == part
          ]

          number <- ifelse(length(number) == 0, 1, number + 1)

          probability <- probability * (number / (author$words + .self$words))
        }

        probability <- probability * author$probability
      
        # find max probability
        if (length(max_probability) == 0 | probability > max_probability) {
          max_probability <- probability
          result_author <- author$author
        }
      }

      return(result_author)
    },

    # score you test set so to get the understanding how well you model
    # works.
    # look at f1 score or precision and recall
    # visualize them
    # try how well your model generalizes to real world data!
    score = function(test_data) {
      predictions <- lapply(test_data$text, function(x) predict(x))

      correct_predictions <- sum(predictions == test_data$author)
      total_observation_count <- nrow(test_data)

      accuracy <- correct_predictions / total_observation_count

      return(accuracy)
    },

    viation = function(test_data) {
      result_df <- data.frame(
        text = character(0),
        author = character(0),
        prediction = character(0),
        result = logical(0)
      )

      for (i in 1:nrow(test_data)) {
        row <- test_data[i, ]
        text <- row$text
        author <- row$author


        prediction <- model$predict(text)

        correct <- ifelse(prediction == author, TRUE, FALSE)

        result_df <- rbind(result_df, data.frame(
          text = text,
          author = author,
          prediction = prediction,
          result = correct
        ))
      }
  
      authors <- unique(test_data$author)
      precision <- numeric()
      recall <- numeric()
      f1_score <- numeric()

      for (author in .self$writers$author) {
        true_positives <- sum(test_data$author == author & result_df$prediction == author)
        false_positives <- sum(test_data$author != author & result_df$prediction == author)
        false_negatives <- sum(test_data$author == author & result_df$prediction != author)

        precision_author <- true_positives / (true_positives + false_positives)
        recall_author <- true_positives / (true_positives + false_negatives)

        precision_author[is.na(precision_author)] <- 0
        recall_author[is.na(recall_author)] <- 0

        if (precision_author + recall_author == 0) {
          f1_score_author <- 0
        } else {
          f1_score_author <- 2 * (precision_author * recall_author) / (precision_author + recall_author)
        }

        precision <- c(precision, precision_author)
        recall <- c(recall, recall_author)
        f1_score <- c(f1_score, f1_score_author)
      }

      avg_precision <- mean(precision, na.rm = TRUE)
      avg_recall <- mean(recall, na.rm = TRUE)
      avg_f1_score <- mean(f1_score, na.rm = TRUE)
      
      cat("Precision:", precision, "\n")
      cat("Recall:", recall, "\n")
      cat("F1 Score:", f1_score, "\n\n")

      cat("Average Precision:", avg_precision, "\n")
      cat("Average Recall:", avg_recall, "\n")
      cat("Average F1 Score:", avg_f1_score, "\n")
      metrics <- data.frame(
        Metric = c("Precision", "Recall", "F1 Score"),
        Value = c(avg_precision, avg_recall, avg_f1_score)
      )

      ggplot(metrics, aes(x = Metric, y = Value, fill = Metric)) +
        geom_bar(stat = "identity") +
        labs(
          title = "Average Precision, Recall, and F1 Score",
          x = "Metric", y = "Value"
        ) +
        theme_minimal()
    }
  )
)
```

```{r}
model <- naiveBayes()
model$fit(tidy_text)
model$predict("I beg both your pardons but I can't be so much mistaken.")
```

```{r}
cat("Accuracy:", model$score(test))
```

## Measure effectiveness of your classifier

Accuracy is not always a good metric for your classifier, so we decided
to use F1 score metric.

$$
{\displaystyle F_{1} Score=2{\frac {\mathrm {precision} \cdot \mathrm {recall} }{\mathrm {precision} +\mathrm {recall} }}}
$$

Where, Precision and Recall equals to

$$
Precision = \frac{true\ positive}{true\ positive\ + \ false \ positive}\\
Recall = \frac{true\ positive}{true\ positive\ + \ false \ negative}\\
$$

```{r}
model$viation(test)
```

## Conclusions

-   In this laboratory we created Naive Bayes Classifier. It uses Bayes
    formula at its core. The Naive Bayes classifier simplifies finding
    probability by assuming feature independence within each class,
    allowing us to calculate the likelihood as the product of individual
    feature probabilities within the class.
-   The pros of this approach is its simplicity, efficiency and good
    performance on small datasets. The cons of Naive Bayes Classifier
    are independence assumption, sensitivity to feature scales and
    limited expressiveness.
